\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-04-11 doc/sw-cfhtls.tex
read through and fixed sect 1-6

In this section, we describe the \sw-\cfhtls lens sample (henceforth,
called the \sw lens sample). We find a total of XXX (and XXX new)
candidates with averaged grade $0<$G$<1.3$ (low probability sample)
which means at least one of the inspector gave a grade of 1 and the
maximum grade given by all three inspectors was 1\footnote{If grades
from the inspectors were found to be discrepant by 2 or more, these were
discussed and re-graded to resolve the discrepancy}. Further information
on the low probability sample such as their positions and images are
available at http://spacewarps.org/\#/projects/CFHTLS/.

There are a total of XXX candidates with $1.3\le$G$<2$ (medium
probability sample) out of which 30 are new. These grades suggest that,
most of the times, at least one of the inspectors gave a grade of 2 and
the maximum grade by all of the inspectors was below 2. Among our high
probability sample (G$\ge$2), there are a total of XXX candidates out of
which 28 are new which means at least two of the inspectors gave a grade
of 2 or higher. To avoid duplication, further information on \sw
candidates that were previously identified in the literature is
available at http://spacewarps.org/\#/projects/CFHTLS/ whereas the new
lens candidate sample of medium-high probability is presented in this
paper (see \Sref{sec:results:newcand}).

In Table~\ref{tab:stats}, we give overall statistics of the systems detected at
\StageOne and \StageTwo. We give the total number of detections of the known lens
candidates, known lenses and the new lens candidates at each stage.  We also
show their fractions with respect to the total number of lens candidates found
at these stages. We note that most (XXX check) of the confirmed lenses found at
\StageOne are also recovered at \StageTwo whereas a small fraction of the known
lens candidates are rejected at \StageTwo. Also, the sample of new lens
candidates increases the sample of \cfhtls lens candidates by over 50\%.

In \Fref{fig:stackP}, we plot the distribution of false positives and the high
probability lens candidates, categorized by the lens experts, as a function of
the P value assigned by SWAP at the end of \StageTwo. On average, the fraction of
lens candidates is indeed an increasing function of P. This shows that the \sw
generated P values for the subject are roughly correlated with the expert grades
albeit with quite some scatter.  We note that from P$\sim$0.75 onwards to lower
P values, the fraction of false positives starts to exceed the fraction of real
lens candidates. This could be a good threshold to choose to maximize the purity
of the final sample. However, the choice of P$=$0.3 as the threshold gives a
more acceptable completeness of XXX for our sample which combined with expert
grading allows us to achieve further purity of our sample.
%{\bf AM: give some numbers for the completeness/purity ? I will check what grades we gave to the stage 2 detected KLs and comment on it.}

@@

We give basic information about the final sample of 58 new lens
candidates (medium-high probability) found by \sw in
Table~\ref{tab:swcands}. We report the candidates with a \sw ID and Name of the
lens system. We give their positions (Ra, Dec), photometric redshift
($z_{phot}$), $i$ band magnitude of the lensing galaxy, averaged grade G from
the lens experts, zoo ID (identifier used in TALK), P
value at \StageTwo and a visual categorization of type of lensed images and the
lensing galaxy in the Comments column in Table~\ref{tab:swcands}. Whenever
available the lens properties are taken from the \cfhtls photometric catalog
\citep{Coupon2009} otherwise the reported lens galaxy positions are measured
manually. The visual categorization of the lens type is only suggestive and the
explanation of the notations in the Comments column is given at the bottom of
the table.

@@

The new \sw lens sample presented here shows the advantages of having
citizen scientists find lenses through visual inspection. An algorithm,
by definition will find objects that adhere to a selection criteria that
uses either geometry or flux information from an image. On the other
hand, citizen scientists can do amazing amount of interpolation or
extrapolation over the basic selection criteria provided to them. 
For example,  the lower blue arc in SW7 is split by a small red galaxy.
An algorithm typically fails to detect such arcs because the arc is
broken into smaller arclets which then falls below the minimum length or
area allowed for an arc to be detected typically. Human brains have no
problem in interpolating over the broken blue arc over the red galaxy
and understand that it is a single long arc. The system SW19 has
point-like lensed images which cannot be detected by arc finding
algorithm whereas the ring finding algorithm may miss this because of
atypical color and structure of the galaxy.

The power of citizen scientists lies in the high dynamic range that
allows us to detect systems with very from small-to-large arcs, from
highly compact-to-low surface brightness images, from round and
point-like to elongated and curved images, from blue-to-red, from
regular-to-exotic kinds of lenses while keeping the false positive rate
comparatively low. Further detailed qualitative and quantitative
analysis of the properties of the entire \sw sample (new and previously
identified candidates) and the mass modelling analyses for the new
candidates will be presented in a subsequent \sw paper, Paper III.

@@

In the subsequent sections, we make comparison of various properties of the lens
candidates. In this section, we describe how we extract these properties, namely,
the lens redshift, the Einstein radii and the total flux of the lensed images or
arcs.

We use the publicly available redshifts for the lens galaxy from the
\cfhtls photometric catalogs \citep{Coupon2009}. The total flux of the
lensed image or arc is measured in the $g$ band but the adopted method
is different for different samples. For the simulated sample, we
multiply the magnification of the second brightest image with the source
magnitude. For the \rf sample, the arcs are detected in the scaled
difference image between $g$ and $i$ bands where the lensing galaxy is
subtracted \citep[for details, see]{Gavazzi2014}. Here, we use the flux
of the lensed images measured by {\sc sextractor} from the scaled
difference image, that is, $g-\alpha i$ and convert it to the $g$ band
flux using mean colors of the foreground and background population. For
the \af and the \sw sample, we integrate the flux in the image pixels
identified by \af or {\sc sextractor}.

The Einstein radius ($R_{\rm E}$) is also measured differently for
different samples. For the galaxy-scale lenses in the simulated sample,
we use the value of the input lens model parameter for the $R_{\rm E}$.
For groups-scale lenses, since the lens model is multi-component, we
need to determine the $R_{\rm E}$ from the image positions. We use those
pairs of lensed images that have the smallest and the largest angular
separations. The $R_{\rm E}$ here is then half of the averaged values of
these angular image separations.  For the \rf sample, we use the peak
position of the lensed images measured by running {\sc sextractor} on
the scaled difference image. We calculate the image separation from the
lens center as an estimate of the $R_{\rm E}$. For the \af (SARCS)
sample, we use the same definition as above except that the peak
position is identified either by the \af or manually. For the \sw lens
sample, the same definition is used where the peak positions are
identified either with \af or {\sc sextractor}.

@@

Indeed, we have find several new lens candidates that were not known
before. In \Fref{fig:isd}, we show the image separation distribution for
the known sample (green), for the \sw identified (known and new) lens
sample (blue) only and combined known with the new \sw lens sample
(magenta). It is interesting to note that both these samples have very
similar profiles and thus, the profile of the combined sample has not
changed much. This implies that previous constraints on the image
separation distribution are robust and the \af selected sample does not
suffer from significant incompleteness for medium to large Einstein
radii. This is the regime that probes density profiles of galaxy groups
to clusters.
%%{\bf AM: understand what Phil wanted to say about AF's (in)completeness}

>>>>

In this section, we describe the \sw-\cfhtls lens sample (henceforth,
called the \sw lens sample). We find a total of XXX (and XXX new)
candidates with averaged grade $0<$G$<1.3$ (low probability sample)
which means - in the worst case, at least one of the inspector thought
the candidate is possibly a lens and in the best case, all three
inspectors thought the candidate is possibly a lens (that is,
a grade of 1)\footnote{If grades from the inspectors were found to be discrepant
by 2 or more, these were discussed and re-graded to resolve the
discrepancy.}. Further information on the low probability sample such as
their positions and images is available at
http://spacewarps.org/\#/projects/CFHTLS/. Note that if all of the inspectors gave
a grade of 0 to a candidate, then it is discarded from the sample.

There are a total of XXX candidates with $1.3\le$G$<2$ (medium
probability sample) out of which 30 are new. The quality of candidates
in this category is such that at least one of the inspectors thought the
candidate is most probably a lens (that is, a grade of 2) and a second inspector
thought this is possibly a lens (that is, a grade of 1). Among our high
probability sample (G$\ge$2), there are a total of XXX candidates out of
which 28 are new. In this category, the minimum grade by all of the
inspectors was 2 suggesting that the candidates are probably or almost
certainly lenses according to all of the inspectors. To avoid
duplication, further information on \sw candidates that were previously
identified in the literature is available at
http://spacewarps.org/\#/projects/CFHTLS/ whereas the sample of new lens
candidates with G$\ge 1.3$ (medium-high probability) is presented in this paper (see
\Sref{sec:results:newcand}).

In Table~\ref{tab:stats}, we give overall statistics of the systems
detected at \StageOne and \StageTwo. We give the total number of
detections of the known lens candidates, known lenses and the new lens
candidates at each stage.  We also give the recovery fractions for the
known samples and fraction of total detection for the new samples at
each stage. We find that most (XXX check) of the confirmed lenses found
at \StageOne are also recovered at \StageTwo whereas a small fraction of
the known lens candidates are rejected at \StageTwo. Also, the sample of
new lens candidates increases the sample of \cfhtls lens candidates by
over 50\%.

In \Fref{fig:stackP}, we plot the distribution of false positives and
the high probability lens candidates, categorized by the lens experts,
as a function of the P value assigned by SWAP at the end of \StageTwo.
On average, the fraction of lens candidates is indeed an increasing
function of P. This shows that the \sw generated P values for the
subject are roughly correlated with the expert grades albeit with quite
some scatter.  We note that from P$\sim$0.75 onwards to lower P values,
the fraction of false positives starts to exceed the fraction of real
lens candidates. This could be a good threshold to choose to maximize
the purity of the final sample. However, we go with the choice of
P$=$0.3 as this gives a more acceptable completeness of XXX for the
known sample and thus, the new sample which combined with expert grading
allows us to achieve further purity of our sample.

%{\bf AM: give some numbers for the completeness/purity ? I will check what grades we gave to the stage 2 detected KLs and comment on it.}

@@

We give basic information about the final sample of 58 new lens
candidates (medium-high probability) found by \sw in
Table~\ref{tab:swcands}. We report the candidates with a \sw ID and Name
of the lens system. We give their positions (Ra, Dec), photometric
redshift ($z_{phot}$), $i$ band magnitude of the lensing galaxy,
averaged grade G from the lens experts, zoo ID (identifier used in
TALK), P value at \StageTwo and a visual categorization of the type of
lensed images and the lensing galaxy in the ``Comments'' column in
Table~\ref{tab:swcands}. Whenever available the lens properties are
taken from the \cfhtls photometric catalog \citep{Coupon2009} otherwise
for the lens galaxy positions, the reported values are measured
manually. The visual categorization of the lens type is only suggestive
and the explanation of the notations in the Comments column is given at
the bottom of the table.

@@

The new \sw lens sample presented here shows the advantages of having
citizen scientists find lenses through visual inspection. An algorithm,
by definition will find objects that adhere to a selection criteria that
uses either geometry or flux information from an image. On the other
hand, citizen scientists can do amazing amount of interpolation or
extrapolation over the basic selection criteria provided to them.  For
example, the lower blue arc in SW7 is split by a small red galaxy.  An
algorithm typically fails to detect such arcs because the arc is broken
into smaller arclets which then falls below the minimum length or area
allowed for an arc to be detected typically. Human brains have no
problem in interpolating over the broken blue arc over the red galaxy
and understand that it is a single long arc. The system SW19 has
point-like lensed images which cannot be detected by arc finding
algorithm whereas the ring finding algorithm may miss this because of
atypical color and structure of the galaxy. Detection of red arcs, for
example, as seen in the SW38 candidate show how the volunteers
extrapolate on the color parameter because the training sample mostly
has blue arcs since this is the typical expectation.

The power of citizen scientists also lies in the high dynamic range that
allows us to find systems with very from short (thick) to long (thin)
arcs, from highly compact to low surface brightness images, from round
and point-like to elongated and curved images, from blue to red, from
regular to exotic kinds of lenses while keeping the false positive rate
low compared to algorithms. Discovery of this large sample of completely
new candidates missed by some of these algorithms demonstrates that the
\sw system is functioning well and the self-taught citizen scientists
have managed to beat the robots in some sense.

Further detailed qualitative and quantitative analysis of the properties
of the entire \sw sample (new and previously identified candidates) and
the mass modelling analyses for the new candidates will be presented in
a subsequent \sw paper, Paper III.

@@

In the subsequent sections, we make comparison of various properties of the lens
candidates. Here, we describe how we extract or measure these properties, namely,
the lens redshift, the Einstein radii and the total flux of the lensed images or
arcs.

We use the publicly available redshifts for the lens galaxy from the
\cfhtls photometric catalogs \citep{Coupon2009}. The total flux of the
lensed image or arc is measured in the $g$ band but the adopted method
is different for different samples. For the simulated sample, we
multiply the magnification of the second brightest image with the source
magnitude. For the \rf sample, the arcs are detected in the scaled
difference image of $g$ and $i$ bands from which the lensing galaxy is
subtracted \citep[for details, see]{Gavazzi2014}. Here, we use the flux
of the lensed images measured by {\sc sextractor} from the scaled
difference image, that is, $g-\alpha i$ and convert it to the $g$ band
flux using mean colors of the foreground and background population. For
the \af and the \sw sample, we integrate the flux in the image pixels
identified by \af or {\sc sextractor}.

The Einstein radius ($R_{\rm E}$) is also measured differently for
different samples. For the galaxy-scale lenses in the simulated sample,
we use the value of the input parameter of the lens model for the $R_{\rm E}$.
For groups-scale lenses, since the lens model is multi-component, we
need to determine the $R_{\rm E}$ from the image positions. We use those
pairs of lensed images that have the smallest and the largest angular
separations. The $R_{\rm E}$ here is then half of the averaged values of
these angular image separations. For the \rf sample, we use the peak
position of the lensed images measured by running {\sc sextractor} on
the scaled difference image. We calculate the image separation from the
lens center as an estimate of the $R_{\rm E}$. For the \af (SARCS)
sample, we use the same definition as above except that the peak
position is identified either by the \af or manually. For the \sw lens
sample, the same definition is used where the peak positions are
identified either with \af or {\sc sextractor}.

@@

Indeed, we have found several new lens candidates that were not known
before. In \Fref{fig:isd}, we show the image separation distribution for
the known \rf and \af sample (green), for the \sw identified (known and
new) lens sample (blue) only and combined \cfhtls sample of \rf, \af and
the new \sw lens sample (magenta). It is interesting to note that both
these samples have very similar profiles and thus, the profile of the
combined sample has not changed signficantly. This implies that previous
constraints on the image separation distribution are robust and the \af
selected sample does not suffer from significant incompleteness for
medium to large Einstein radii. This is the regime that probes density
profiles of galaxy groups to clusters.
%%{\bf AM: understand what Phil wanted to say about AF's (in)completeness}
